\bf An HMM-based part-of-speech tagger for Turkish

### 2015-12-07 paper/paper.tex
New sample and arrows in the HTML table
Write about performance

In general, most of the literature relating to the computational processing of Turkish
pertain primarily to morphology---compared to the complexity of Turkish
morphology, its syntax seems to be a less interesting problem.
\section{Proposal}

@@

>>>>

In general, most of the literature relating to the computational processing of Turkish
pertain primarily to morphology---compared to the complexity of Turkish
morphology, its syntax seems to be a less interesting problem.
\section{Proposed Work}

@@

\section{Results}
The POS tagger I've built as proposed can successfully tag words with a
reasonable accuracy.

\subsection{Performance}
Parsing the entire corpus and creating a model out of it takes 10.5 seconds on
average. Tagging a sentence (which we're assuming to have 5 words on average)
takes $~0.3$ seconds. The implementation of the \texttt{Data.HMM} datatype uses
\texttt{Data.Array}'s
\footnote{https://hackage.haskell.org/package/array-0.5.1.0/docs/Data-Array.html}
which is why it runs reasonably fast.

\subsection{Limitations}
There is a fundamental limitation to the POS tagger I built: Turkish morphology
is essential to the overall completeness of Turkish.

In English we entertain the notion of \emph{syntactic correctness}, which is
more or less a formalism of correct ordering of words. A speaker of English that has
synactic competence in the language \citep{bachman1990} may know that a sentence
is well-formed, relying on his knowledge of syntactic rules. However, for the
\textit{words}, knowledge of morphological rules is not adequate to conclude
that a word is made up of correctly ordered morphemes; they have to rely on
their lexical knowledge to verify that the morphological form they
derived/inflected is a valid one. Consider the word ``satisfiableness'' which
is not a valid word according to several dictionaries, although being derived
with valid suffixes. The correct form of this concept is ``satisfiability''.

In Turkish, the concept of word validness is a lot like syntactic correctness;
the speaker may rely their knowledge of morphological rules to \emph{generate}
new words. In fact Turkish has \textbf{recursive} morphology \citep{kabak2001}
so a word can be inflected/derived in into infinitely many different forms. This
is why computational processing of Turkish can never fully escape its
morphology.

The hidden markov model I've built for this project \emph{ignores} morphology.

\section{Conclusion}

% Delete this todo
\todo[inline]{Write a concluding summary. Also indicate what future work may be warranted. This section should be at least half a page in length.}
